---
title: "FA6_KHAFAJI"
author: "Mostafa Khafaji"
date: "2024-02-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## 1. Geometric Distribution
**rovide an R code for the geometric distribution. The geometric distribution is a probability distribution that models the number of trials required to achieve the first success in a sequence of Bernoulli trials, where each trial has a constant probability of success.**

1. Set the probability of success:  p <- 0.2

2. Generate 1000 random variables from the geometric distribution.

3. Calculate some basic statistics:

  mean_x <- mean(x)

  var_x <- var(x)

  sd_x <- sd(x)
  
4. Print the results in item 3 with the following output (string): 

  Number of trials required to achieve first success:

  Mean (in 2 decimal places):

  Variance (in 2 decimal places):

  Standard deviation ( in 2 decimal places):
  
```{r geom_descrips, echo=TRUE}
#1
p <- 0.2

#2
k <- 1000
x <- rgeom(k, prob=p)

#3
mean_x <- mean(x)
var_x <- var(x)
sd_x <- sd(x)

#4
cat("The minimum number of trials to get a successful event is:", min(x)+1,"\n\n")
cat("The mean (in 2 decimal places) is: ",sprintf(mean_x, fmt='%#.2f'),"\n\n")
cat("The variance (in 2 decimal places) is: ",sprintf(var_x, fmt='%#.2f'),"\n\n")
cat("The Standard Deviation (in 2 decimal places) is: ",sprintf(sd_x, fmt='%#.2f'),"\n\n")

```

5. Plot the histogram of the results.

```{r}

hist(x, main = "Number of Failures before Success", xlab="Number of Failures", breaks = 25)
```


## 2. Hypergeometric Distribution.  

**Consider a plant manufacturing IC chips of which 10% are expected to be defective. The chips are packed in boxes for export. Before transportation, a sample is drawn from each box. Estimate the probability that the sample contains more than 10% defectives, when:**


  1. A sample of 10 is selected from a box of 40;
  
```{r hgeom_1, echo=TRUE}
#setting variables
Popsize1 <- 40
pop_ex_failures1 <- 40*0.1
sample_size <- 10
#since more than 10%, we add 1
samp_ex_failures <- as.integer(sample_size*0.1)+1

#calculating probability
prob1 <- (choose(pop_ex_failures1, samp_ex_failures))*(choose(Popsize1 - pop_ex_failures1, sample_size - samp_ex_failures))/choose(Popsize1,sample_size)

cat("The probability that more than 10% of a sample of 10 picked out from 40 is a failure is: ",prob1*100,"%",sep = "")

```
  
  2. A sample of 10 is selected from a box of 5000.

```{r hgeom_2, echo=TRUE}
#setting variables
popsize2 <- 5000
pop_ex_failures2 <- popsize2*0.1

#calculating probability
prob2 <- (choose(pop_ex_failures2, samp_ex_failures))*(choose(popsize2 - pop_ex_failures2, sample_size - samp_ex_failures))/choose(popsize2,sample_size)

cat("The probability that more than 10% of a sample of 10 picked out from 5000 is a failure is: ",prob2*100,"%",sep = "")
```

